---
title: "Time Series Analysis on Crime Data"
author: '232806010,232807013'
date: "2024-04-25"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

### Importing libraries required

```{r}

library(readxl)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tseries)
library(forecast)
library(MASS)
```

### Importing the data set

```{r}
data=read.csv("merged_crime_data.csv")
head(data)
```

### Grouping the data by date

```{r}
crime_data<- data %>%
  mutate(Date = as.Date(Date.Occurred)) %>%
  group_by(Date) %>%
  summarise(crime_count = n())
head(crime_data)
```

### Re-sampling to monthly average crime

```{r}
monthly_data<-crime_data %>%
  group_by(month = floor_date(Date, "month")) %>%
  summarise(avg_crime = mean(crime_count, na.rm = TRUE))
head(monthly_data)
```

### Grouping data annually

```{r}
year_data <- monthly_data%>%
  group_by(year = year(month)) %>%
  summarise()
head(year_data)
```

### Creating an empty data frame with the necessary column names and assigning values to the corresponding columns

```{r}
# Create an empty data frame with the necessary column names
annual_crime <- data.frame(matrix(ncol = nrow(year_data), nrow = 12))
colnames(annual_crime) <- as.character(year_data$year)

# Loop through the year_groups
for (i in 1:nrow(year_groups)) {
  year <- year_groups$year[i]          # Pulling the year from datetime object
  if (year == 2024) {
    break
  }
  series <- monthly_crime$avg_crime[year(monthly_crime$month) == year]  # Extracting data for the year
  # Assign values to the corresponding column
  annual_crime[, as.character(year)] <- series
}
head(annual_crime)
```

```{r}
# Boxplot for last 10 years
boxplot(annual_crime, main = "Los Angeles Volume of Crime Trend between 2010 - 2024",
        xlab = "Years", ylab = "Monthly Average Crimes Reported", col = "lightblue", outline = FALSE)

```
### Converting the data into a time series data

```{r}
monthly_data_ts <- ts(monthly_data$avg_crime, start = c(year(monthly_data$month)[1], month(monthly_data$month)[1]), frequency = 12)
monthly_data_ts
```

### Exploratory Data Analysis

### Time series plot of crimes by day of week

```{r}
data$`Day of Week` <- wday(data$`Date.Occurred`, label = TRUE)
data$`Day of Week` <- factor(data$`Day of Week`, levels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))
ggplot(data, aes(x = `Day of Week`)) +
  geom_bar(fill = "lightgreen", color = "black") +
  labs(title = "Number of Crimes by Day of Week",
       x = "Day of Week",
       y = "Number of Crimes") +
  scale_x_discrete(labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))
```

### Plotting categorical count plot for crime by area

```{r}
ggplot(data, aes(x = reorder(`Area.Name`, -table(`Area.Name`)[`Area.Name`]), fill = `Area.Name`)) +
  geom_bar() +
  labs(title = "Los Angeles - Volume of Crime across the 21 LA Geographic Areas",
       x = "Volume of Crime in an Area",
       y = "Area Name") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10))  # Adjust font size for y-axis labels

```

### Boxplot

### Visualization of time series patterns

```{r}
autoplot(decompose(monthly_crime_ts))
```

### ACF Plot

```{r}
Acf(monthly_crime_ts, main = "Autocorrelation Function (ACF) for Average Crime")
```

### PACF Plot

```{r}
Pacf(monthly_crime_ts, main = "Partial Autocorrelation Function (PACF) for Average Crime")
```

### Splitting the data set into train and test data

```{r}
#train data 
crime_train=ts(monthly_data_ts,start=c(2010,1),end=c(2019,12),frequency = 12)
crime_train 

```

```{r}
#test data 
crime_test=ts(monthly_data_ts,start=c(2020,1),end=c(2024,4),frequency=12)
crime_test
```

### Checking for stationarity

```{r}
adf_test=adf.test(crime_train)
adf_test
#data is not stationary 
```

### Box-Cox Plot

```{r}
boxcox_result=boxcox(crime_train~1)
boxcox_result
plot(boxcox_result)
```

### To find the lambda value

```{r}
lambda=boxcox_result$x[which.max(boxcox_result$y)]
lambda
```

### Transforming the data using log transformation

```{r}
train_log=log(crime_train)
test_log=log(crime_test)
```

### Checking for stationarity

```{r}
adf_test=adf.test(train_log)
adf_test # data is not stationary
```

### Differencing the data and then checking the stationarity

```{r}
train_log_diff=diff(train_log)
adf_test=adf.test(train_log_diff)
adf_test # data is stationary
```

### To find the best ARIMA model for the log transformed data

```{r}
auto.arima(train_log)
```

### Fitting a SARIMA model on the log transformed data

```{r}
sarima_log=arima(train_log,order=c(p=1,d=1,q=3),seasonal=c(P=2,D=0,Q=0))
sarima_log #AIC=-470.88
```

### Using the SARIMA model to predict the average crime for the next 52 months

```{r}
predicted_log_sarima=forecast(sarima_log,h=52)
predicted_log_sarima
```

### Plotting the predicted values

```{r}
autoplot(predicted_log_sarima)
```

### Checking the accuracy

```{r}
accuracy(predicted_log_sarima,test_log)
```

### Checking the residuals

```{r}
checkresiduals(sarima_log)
```

### Checking the normality of the residuals

```{r}
shapiro.test(residuals(sarima_log))#not normal 
```

### Checking for autocorrelation between the residuals

```{r}
Box.test(residuals(sarima_log))#residuals are not autocorrelated
```

### Fitting a Holt's Winter model on the log transformed data

```{r}
holts_log=hw(train_log,h=52)
summary(holts_log)#AIC=-229.5010
```

### Using the Holt's Winter model to predict the average crime for the next 52 months

```{r}
predicted_log_holts=forecast(holts_log)
predicted_log_holts
```

### Plotting the predicted values

```{r}
autoplot(predicted_log_holts)
```

### Checking the accuracy

```{r}
accuracy(predicted_log_holts,test_log)
```

### Checking the residuals

```{r}
checkresiduals(holts_log)
```

### Checking for normality

```{r}
shapiro.test(residuals(holts_log))#Not notmal
```

### Checking for autocorrelation between the residuals

```{r}
Box.test(residuals(holts_log))#residuals are not autocorrelated 
```

### 

### Transforming the data using square root transformation

```{r}
train_sqrt=sqrt(crime_train) 
test_sqrt=sqrt(crime_test)
```

### Checking for stationarity

```{r}
adf_test=adf.test(train_sqrt) 
adf_test # data is not stationary
```

### Differencing the data and then checking the stationarity

```{r}
train_sqrt_diff=diff(train_sqrt) 
adf_test=adf.test(train_sqrt_diff) 
adf_test # data is stationary
```

### To find the best ARIMA model for the square root transformed data

```{r}
auto.arima(train_sqrt)
```

### Fitting a SARIMA model on the square root transformed data

```{r}
sarima_sqrt=arima(train_sqrt,order=c(p=1,d=1,q=3),seasonal=c(P=2,D=0,Q=0))
sarima_sqrt #AIC=119.58
```

### Using the SARIMA model to predict the average crime for the next 52 months

```{r}
predicted_sqrt_sarima=forecast(sarima_sqrt,h=52)
predicted_sqrt_sarima
```

### Plotting the predicted values

```{r}
autoplot(predicted_sqrt_sarima)
```

### Checking the accuracy

```{r}
accuracy(predicted_sqrt_sarima,test_sqrt)
```

### Checking the residuals

```{r}
checkresiduals(sarima_sqrt)
```

### Checking the normality of the residuals

```{r}
shapiro.test(residuals(sarima_sqrt))#not normal 
```

### Checking for autocorrelation between the residuals

```{r}
Box.test(residuals(sarima_sqrt))#residuals are not autocorrelated
```

### Fitting a Holt's Winter model on the square root transformed data

```{r}
holts_sqrt=hw(train_sqrt,h=52)
summary(holts_sqrt)#AIC=371.7864
```

### Using the Holt's Winter model to predict the average crime for the next 52 months

```{r}
predicted_sqrt_holts=forecast(holts_sqrt) 
predicted_sqrt_holts
```

### Plotting the predicted values

```{r}
autoplot(predicted_sqrt_holts)
```

### Checking the accuracy

```{r}
accuracy(predicted_sqrt_holts,test_sqrt)
```

### Checking the residuals

```{r}
checkresiduals(holts_sqrt)
```

### Checking for normality

```{r}
shapiro.test(residuals(holts_sqrt))#Not notmal
```

### Checking for autocorrelation between the residuals

```{r}
Box.test(residuals(holts_sqrt))#residuals are not autocorrelated 
```

### Transforming the data using reciprocal transformation

```{r}
train_rec=(1/(crime_train))
test_rec=(1/(crime_test))
```

### Checking for stationarity

```{r}
adf_test=adf.test(train_rec)  
adf_test # data is not stationary
```

### Differencing the data and then checking the stationarity

```{r}
train_rec_diff=diff(train_rec)  
adf_test=adf.test(train_rec_diff)  
adf_test # data is stationary
```

### To find the best ARIMA model for the reciprocal transformed data

```{r}
auto.arima(train_rec)
```

### Fitting a SARIMA model on the reciprocal transformed data

```{r}
sarima_rec=arima(train_rec,order=c(p=0,d=1,q=2),seasonal=c(P=2,D=0,Q=0)) 
sarima_rec #AIC=-1980.74
```

### Using the SARIMA model to predict the average crime for the next 52 months

```{r}
predicted_rec_sarima=forecast(sarima_rec,h=52)
predicted_rec_sarima
```

### Plotting the predicted values

```{r}
autoplot(predicted_rec_sarima)
```

### Checking the accuracy

```{r}
accuracy(predicted_rec_sarima,test_rec)
```

### Checking the residuals

```{r}
checkresiduals(sarima_rec)
```

### Checking the normality of the residuals

```{r}
shapiro.test(residuals(sarima_rec))#not normal 
```

### Checking for autocorrelation between the residuals

```{r}
Box.test(residuals(sarima_rec))#residuals are not autocorrelated
```

### Fitting a Holt's Winter model on the reciprocal transformed data

```{r}
holts_rec=hw(train_rec,h=52) 
summary(holts_rec)#AIC=-1626.289
```

### Using the Holt's Winter model to predict the average crime for the next 52 months

```{r}
predicted_rec_holts=forecast(holts_rec)  
predicted_rec_holts
```

### Plotting the predicted values

```{r}
autoplot(predicted_rec_holts)
```

### Checking the accuracy

```{r}
accuracy(predicted_rec_holts,test_sqrt)
```

### Checking the residuals

```{r}
checkresiduals(holts_rec)
```

### Checking for normality

```{r}
shapiro.test(residuals(holts_rec))#residuals are  normal
```

### Checking for autocorrelation between the residuals

```{r}
Box.test(residuals(holts_rec))#residuals are autocorrelated 
```

### Using the SARIMA model on the data transformed using reciprocal transformation

### Forecast for the next the next 72 months

```{r}
model=arima(train_rec,order=c(p=0,d=1,q=2),seasonal=c(P=2,D=0,Q=0))
model
```

```{r}
forecast_model=forecast(model,h=72)

```

### Plotting the forecasts

```{r}
autoplot(forecast_model)
```
